---
title: "Insert paper title here"
author: "Ada Yan"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{main}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

This vignette contains code to reproduce the modelling analysis and figures in the manuscript "", using data and functions in the `deltaomicron1` package.

## Main text

```{r global_options, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

```{r}
library(deltaomicron1)
```

```{r eval = TRUE, include = FALSE}
devtools::load_all()
```

`run_exp_camostat_amphoB` is the function used to fit the model to the data.  For example,

```{r eval = FALSE}
run_exp_camostat_amphoB(save_dir = "results/",
                        strain = "omicron",
                        camostat_vec = c(FALSE, TRUE, FALSE),
                        amphoB_vec = c(FALSE, FALSE, TRUE),
                        Calu3 = TRUE,
                        PCR = TRUE,
                        mvr = FALSE,
                        length_run = 2,
                        run_flag = TRUE)
```

fits the model to the Omicron Calu-3 data without drugs and with one drug, and saves the results in results/.  

The arguments `camostat_vec = c(FALSE, TRUE, FALSE)` and `amphoB_vec = c(FALSE, FALSE, TRUE)` indicate the data sets to be fitted to.  Here we fit to three data sets, corresponding to each value of the vectors `camostat_vec` and `amphoB_vec`.  The first values of `camostat_vec` and `amphoB_vec` are `FALSE` and `FALSE`, indicating the no-drug data set; the second values are `TRUE` and `FALSE` respectively, indicating the Camostat-only data set; the final values are `FALSE` and `TRUE` respectively, indicating the Amphotericin B-only data set.

The remaining arguments to this function are:

* `save_dir`: directory to save results to
* `strain`: `delta` or `omicron`: virus strain of data set to fit to.  (`omicron` is BA.1)
* `Calu3`: if `TRUE`, fit to Calu-3 cell data, if `FALSE`, fit to hNEC data
* `PCR`: if `TRUE`, fit to qPCR and plaque assay data simultaneously; if `FALSE`, fit to plaque assay data only
* `mvr`: if `TRUE`, use multivariate proposal distribution for MCMC; if `FALSE`, use univariate proposal distribution for MCMC
* `length_run`: values `1`, `2` and `3` are used for running short, mid-length, or long MCMC chains respectively
* `run_flag`: if `TRUE`, run the MCMC sampler, if `FALSE`, only perform postprocessing

Most of these arguments take a single value in this work, because the scope of the code is greater than the scope of the work.

To run all of the model fits in the main text:

```{r eval = FALSE}
input_grid <- expand_grid(strain = c("omicron", "delta"), Calu3 = c(TRUE, FALSE)) %>%
  mutate(output_folder = paste0(strain, "_", Calu3, "/"))
apply_named_args(input_grid, 1, function(strain, Calu3, output_folder) run_exp_camostat_amphoB(output_folder,
                                                                          strain = strain,
                                                                          camostat_vec = c(FALSE, TRUE, FALSE),
                                                                          amphoB_vec = c(FALSE, FALSE, TRUE),
                                                                          Calu3 = Calu3,
                                                                          PCR = TRUE,
                                                                          mvr = FALSE,
                                                                          length_run = 2,
                                                                          run_flag = TRUE))
```

The main output of the code are the files `1_chain.csv`, `6_chain.csv`, and `11_chain.csv`, which contain the parameter values sampled by the MCMC chains.  3 MCMC chains are run for each model fit, with one output file corresponding to each.  Additionally, summary statistics are computed, and their median and 95\% credible intervals, as well as those of the raw parameters, are stored in `1_prctile.tex`, `6_prctile.tex` and `11_prctile.tex` for each of the three chains.  If the three MCMC chains have converged, then there is an additional file `1_prctile_combined.tex` which contains the median and 95\% credible intervals combined across the three chains.

To make the model predictions and plots in Fig. X:

```{r eval = FALSE}
calc_and_plot_trajectories <- function(strain, Calu3, output_folder) {
  
  trajectory_filename <- paste0(output_folder, "trajectories.rds")
  trajectories <- calc_trajectories_camostat_amphoB(filename = paste0(output_folder, "1.RData"))
  saveRDS(trajectories, trajectory_filename)
  
  plot_trajectories(strain, Calu3 = Calu3, trajectory_filename)
}

apply_named_args(input_grid, 1, calc_and_plot_trajectories)
```

To calculate the p-values for the doubling time differing between Omicron and Delta:

```{r eval = FALSE}
calc_p_values <- function(input_grid, Calu3_in) {
  dir_names <- input_grid %>% filter(Calu3 == Calu3_in) %>% pull(output_folder)
  pathways <- c("tmprss2", "endosomal", "both", "endosomal_no_IFITM", "both_no_IFITM")
  pathways <- paste0("doubling_time", pathways)
  p_values <- lapply(pathways, pairwise_stats_tests, data_dir = dir_names)
  p_values <- vnapply(p_values, function(x) x$p_value)
  names(p_values) <- pathways
  p_values
}
lapply(c(FALSE, TRUE), calc_p_values, input_grid = input_grid)
```